---
title: "Data Project 3: Youtube Activity & Trending Videos Analysis"
output: html_notebook
---

What makes a Youtube video go viral and how does that relate to what makes a Youtube video one that I would like?

Data Sources: Google Takeout, Youtube API

Exploratory Questions:
- What types of videos do I watch? Which youtubers do I watch the most?
- How does what I watch compare with what other people are watching (trending videos)?
- How has my watch/search history and trending videos changed throughout the years? 
- How much time do I spend on Youtube? Are there certain hours of the day I watch more often?

Analysis
- Sentiment Analysis: analyzing sentiment of descriptions and comments on different types of videos
- Predictive Modeling: using category, sentiment, time of day, video length, etc. to predict video’s views, likes, comments




```{r}
library(stringr)
library(rvest) 
library(jsonlite) 
library(tidytext)
library(lubridate) 
library(wordcloud)
library(httr)
library(ggplot2)

library(stringr)
library(dplyr)
library(magrittr)
library(readr)
library(tidyr)
```
 
```{r}
# READ SEARCH HISTORY
search_history = read_html("Takeout/YouTube and YouTube Music/history/search-history.html")
search = search_history %>%
  html_nodes(".header-cell + .content-cell > a") %>%
  html_text()
search_content <- search_history %>%
  html_nodes(".header-cell + .content-cell")
search_time = str_match(search_content, "<br>(.*?)</div>")[,2]
search_time = mdy_hms(search_time)

# CREATING DATA FRAME SEARCH + TIMESTAMP
search_df = data.frame(search = search, 
                                time = search_time,
                                stringsAsFactors = FALSE)
search_df
```


```{r}
library(tm)
corp.original = VCorpus(VectorSource(search_df$search)) 
corp = tm_map(corp.original, removePunctuation)
corp = tm_map(corp, removeNumbers)
corp = tm_map(corp, content_transformer(tolower), lazy=TRUE) 
#corp = tm_map(corp, content_transformer(stemDocument), lazy=TRUE) 
corp = tm_map(corp, content_transformer(removeWords), c("TIL") ,lazy=TRUE) 
corp = tm_map(corp, stripWhitespace)
corp = tm_map(corp, content_transformer(removeWords), c(stopwords("english")), lazy=TRUE)
wordcloud(corp, colors=brewer.pal(11, "Set2"))
```
 
```{r}
# READ WATCH HISTORY 
watch_history <- read_html("Takeout/YouTube and YouTube Music/history/watch-history.html")
watched_videocontent <-  watchHistory %>%
  html_nodes(".header-cell + .content-cell")
# POSSIBLE TIME CHARACTERS
watch_time = str_match(watched_videocontent, 
                             "<br>([A-Z].*)</div>")[,2]
# POSSIBLE ID VALUES 
watched_id = str_match(watched_videocontent, 
                             "watch\\?v=([a-zA-Z0-9-_]*)")[,2]
# VIDEO TITLE
watched_titles = str_match(watched_videocontent, 
                                "watch\\?v=[a-zA-Z0-9-_]*\">(.*?)</a>")[,2]

# DATA FRAME WATCH HISTORY
watched_df = data.frame(id = watched_id, 
                                     title = watched_titles , 
                                     time = watch_time, 
                                     stringsAsFactors = FALSE)
watched_df$time = mdy_hms(watched_df$time)

watched_df = filter(watched_df, watched_df$time > '2021-01-01')
watched_df = filter(watched_df, watched_df$time < '2021-05-01')
watched_df
```

Data goes from December 12, 2020 - May 2, 2021. For consistency between months I filter out the December and May data, leaving 4 full months of data from January to April 2021.

Preliminary Analysis of Watch Frequency
```{r}
#group by month
watched_df$month = as.numeric(month(watched_df$time))

watchbymonth <- watched_df %>%
  group_by(month) %>%
  tally()

barplot(watchbymonth$n,ylab="Watch Count",main="Monthly Watch Frequency", names.arg=watchbymonth$month, col = "blue", las=1, cex.axis=.5, cex.names=1)

#group by day
watched_df$date = format(as.POSIXct(watched_df$time,format='%Y-%m-%d %H:%M:%S'),format='%m-%d-%Y')

watchbydate <- watched_df %>%
  group_by(date) %>%
  tally()
watchbydate$date = as.Date(watchbydate$date, "%m-%d-%Y")

#plot watch frequency these past 4 months
ggplot(data=watchbydate, aes(x=date, y=n, group=1)) +
  geom_line()+
  geom_point() + ylab("Watch Count") + ggtitle("My Youtube Activity Jan-Apr 2021") + xlab("Date")

```
Watch Timing
```{r}
#group by hour of day
watched_df$hour = as.numeric(hour(watched_df$time))

watchbyhour = watched_df %>%
  group_by(hour) %>%
  tally()

barplot(watchbyhour$n,ylab="Watch Count",main="Hourly Watch Frequency", names.arg=watchbyhour$hour, col = "pink", las=1, cex.axis=.5, cex.names=.5)


ggplot(watchbyhour, aes(x = hour, y=n)) +
  geom_bar(width = 2, colour="grey",stat = "identity") +
  theme_minimal() +
  scale_fill_brewer() +
  coord_polar(start=0) +
  scale_x_continuous("", limits = c(0, 24), breaks = seq(0, 24), labels = seq(0,24)) + ggtitle("Watch Frequency by Time of day")
```

Most Rewatched
```{r}
install.packages("kableExtra")
library(kableExtra)


rewatched = watched_df %>%
  group_by(id, title) %>% 
  tally()

rewatched = rewatched %>%
  arrange(desc(n))
rewatched = subset(rewatched, select = c(title,n)) 

names(rewatched)[names(rewatched) == "n"] <- "watch count"

rewatched_top3 = head(rewatched,3)

rewatched_top3 %>%
  kbl(caption = "Top 3 Rewatched Youtube Videos") %>%
  kable_material(c("striped", "hover"))
  
```

Video Preferences

```{r}
install.packages("tuber")
install.packages("httpuv")
library(tuber)
library(httpuv)

client_id = "1097085864260-i4hlng05dle65dfj4bllrombicngk1fd.apps.googleusercontent.com"
client_secret = "T5xi7YFn8Vq3yfixD6w_8vrF"

yt_oauth(app_id = client_id,
         app_secret = client_secret,
         token = '')
```
 
```{r}
get_video_details(video_id="hLZX1gOF_64")

unique_watched = unique(watched_df$id)

get_all_stats <- function(id) {
  get_stats(id)
} 

# Get stats and convert results to data frame 
metadata = lapply(unique_watched, get_all_stats)
metadata_df = do.call(rbind.fill, lapply(res, data.frame))
metadata_df

#
watched_df = merge(x = watched_df, y = rewatched, by = "title")
watched_df = merge(x = watched_df, y = metadata_df, by = "id")

head(watched_df)
```
 
 Youtube Trending Analysis:
 
 Trending vids data collected from Kaggle
```{r}

yt_trending = read_csv("UStrending.csv")
yt_trending

names(yt_trending)[names(yt_trending) == "video_id"] <- "id"

#filter out only jan-apr 2021
yt_trending = filter(yt_trending, yt_trending$trending_date > '2021-01-01')
yt_trending 

```

```{r}
library(plyr)
library(dplyr)

yt_trending$categoryId[which(yt_trending$categoryId == "24")] <- "Entertainment"
yt_trending$categoryId[which(yt_trending$categoryId == "20")] <- "Gaming"
yt_trending$categoryId[which(yt_trending$categoryId == "2")] <- "Autos & Vehicles"
yt_trending$categoryId[which(yt_trending$categoryId == "23")] <- "Comedy"
yt_trending$categoryId[which(yt_trending$categoryId == "10")] <- "Music"
yt_trending$categoryId[which(yt_trending$categoryId == "2")] <- "Autos & Vehicles"
yt_trending$categoryId[which(yt_trending$categoryId == "17")] <- "Sports"
yt_trending$categoryId[which(yt_trending$categoryId == "25")] <- "News & Politics"
yt_trending$categoryId[which(yt_trending$categoryId == "26")] <- "Howto & Style"
yt_trending$categoryId[which(yt_trending$categoryId == "28")] <- "Science & Technology"
yt_trending$categoryId[which(yt_trending$categoryId == "22")] <- "People & Blogs"
yt_trending$categoryId[which(yt_trending$categoryId == "1")] <- "Film & Animation"
yt_trending$categoryId[which(yt_trending$categoryId == "15")] <- "Pets & Animals"
yt_trending$categoryId[which(yt_trending$categoryId == "27")] <- "Education"
yt_trending$categoryId[which(yt_trending$categoryId == "29")] <- "Nonprofits & Activism"
yt_trending$categoryId[which(yt_trending$categoryId == "19")] <- "Travel & Events"


names(yt_trending)[names(yt_trending) == "categoryId"] <- "category"
```


Trending Videos I Watched
```{r}
#filter unique trending videos & # of times trending 
unique_trending  = yt_trending  %>%
  group_by(id, title, category, trending_date) %>% 
  tally()
unique_trending = unique_trending %>%
  arrange(desc(n))

names(unique_trending)[names(unique_trending) == "n"] <- "trending count"

#filter unique watched videos
unique_watched_df  = watched_df %>%
  group_by(id, title) %>% 
  tally()
unique_watched_df
names(unique_watched_df)[names(unique_watched_df) == "n"] <- "watch count"

overlap_df = merge(x = unique_watched_df, y = unique_trending, by = "id")
overlap_df 
```

Sentiment of video description
```{r}
library(tidytext)
library(textdata)

yt_trending_sent = yt_trending  %>% 
  select(id, title, description) %>% 
  unnest_tokens(word, description) %>%        
  inner_join(get_sentiments("afinn")) %>%        
  group_by(id) %>%      
  summarise(emotion = mean(value))
yt_trending_sent

#merge with main df
yt_trending = merge(x = yt_trending, y = yt_trending_sent, by = "id")

#sentiment based on category
cat_sent  = yt_trending %>%
  group_by(category) %>% 
  summarise(avg_sent = mean(emotion))

cat_sent = cat_sent %>% arrange(desc(avg_sent))
cat_sent

yt_trending
#sentiment over time
sent_over_time = yt_trending %>%
  group_by(trending_date, emotion) %>%
  tally()

sent_over_time = sent_over_time %>%
  group_by(trending_date) %>%
  summarise(daily_sent = mean(emotion))
sent_over_time

ggplot(data=sent_over_time, aes(x=trending_date, y=daily_sent, group=1)) +
  geom_line()+
  geom_point() + ylab("Daily Sentiment Score") + ggtitle("Public Sentiment From Youtube Trending") + xlab("Date")


```

Seeing what videos are in end of march
```{r}
yt_trending_mar = filter(yt_trending, yt_trending$trending_date > '2021-03-23')
yt_trending_mar = filter(yt_trending_mar, yt_trending_mar$trending_date < '2021-04-01')
yt_trending_mar = yt_trending_mar %>% arrange(emotion)
yt_trending_mar
```


Correlation between views, likes, dislikes, comments, sentiment
```{r}
#install.packages("reshape2")
library(reshape2)

filter_corr =  yt_trending[, c('likes', 'dislikes', 'comment_count', 'view_count')]
corr_mat <- cor(filter_corr )
round(corr_mat, 2)

melted_cormat = melt(corr_mat)
melted_cormat

# Get upper triangle of the correlation matrix
get_upper_tri <- function(corr_mat){
  corr_mat[lower.tri(corr_mat)]<- NA
  return(corr_mat)
}
  
upper_tri <- get_upper_tri(corr_mat)

#corr matrix
melted_cormat = melt(upper_tri, na.rm = TRUE)
# Heatmap

ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
```

Overlap Trending Over Time
```{r}

overlap_cat = overlap_df %>% 
  group_by(category) %>% 
  tally() 

overlap_cat = overlap_cat %>%
  arrange(desc(n))

overlap_cat

overlap_df %>% 
  ggplot(aes(x = trending_date, fill = category)) + 
  labs(x= "Date", y= "Count") + 
  ggtitle("Trending Videos 2021 Q1 I Watched")+
  geom_area(stat = "bin")
```



Trending Over Time by Category
```{r}
#install.packages("ggthemes")
library(ggthemes)

trending_cat = yt_trending %>% 
  group_by(category) %>% 
  tally() 

trending_cat = trending_cat %>%
  arrange(desc(n))

trending_cat

yt_trending %>% 
  ggplot(aes(x = trending_date, fill = category)) + 
 labs(x= "Date", y= "Count") + 
  ggtitle("Trending Videos 2021 Q1", "Most trending categories")+
  geom_area(stat = "bin")
```
 
 